---
title: "Cleaning the Data"
author: "Charles Marks"
date: "November 30, 2018"
output: html_document
---

## Readme

This file contains the code used to first upload and clean the data for use in analysis.  This doc only contains cleaning, not doing exclusions in this doc

```{r setup, include=FALSE}
library(tidyverse)
library(icd)
library(varhandle)
```

## First, Uploading the Data to R

We have three data tables that contain our entire dataset, so we first need to upload these datasets into R so we can work with them.  Due to some issues with uploading xlsx files, the three xlsx files were saved as csvs (where each sheet represents its own sheet)

```{r upload}


demographics <- read.csv("Demographics.csv", header = TRUE)
medications <- read.csv("Medications.csv", header = TRUE)
hospitalizations <- read.csv("Hospitalization.csv", header = TRUE)
social_hx <- read.csv("Social_Hx.csv", header = TRUE)
vitals <- read.csv("Vitals.csv", header = TRUE)
diagnoses <- read.csv("Diagnoses.csv", header = TRUE)

hospital_billing_CPT_procs <- read.csv("Hospital_Billing_Procs.csv", header = TRUE)
prof_billing_CPT_procs <- read.csv("Prof_Billing_CPT_Procs.csv", header = TRUE)
surgeries <- read.csv("Surgeries.csv", header = TRUE)
orders_procedures <- read.csv("Orders_Procedures.csv", header = TRUE)
labs <- read.csv("Labs.csv", header = TRUE)

medication_class <- read.csv("Medication_Thera_Class.csv", header = TRUE)
rxnorm <- read.csv("RxNorm.csv", header = TRUE)
  

```

## Start Generating the Final Dataset

```{r pressure, echo=FALSE}

# We will use the demographics table as our baseline data table, as in theory, it contains all of the subjects
final_dataset <- demographics
#for some reason the ids in this table were factored, so need to unfactor
final_dataset$MRN <- as.numeric(as.character(final_dataset$MRN))
#now remove all those with no IDs
final_dataset <- final_dataset[!(is.na(final_dataset$MRN)),]
```
## Primary Outcome -- Any Glaucoma Surgery

### Surgeries

So we will go through the surgeries tables and create a series of variables related to surgeries

```{r Surgeries}

## we are going to first fix the date field of the surgery table
surgeries$surgery_date <- as.Date(as.character.Date(surgeries$surgery_date),"%m/%d/%y")

## we will sort the surgeries by date for convenience
surgeries <- surgeries[order(surgeries$surgery_date),]

## we are going to create several surgery variables

final_dataset$any_glaucoma_surgey <- 0
final_dataset$incisional_glaucoma_surgery <- 0
final_dataset$min_invasive_glaucoma_surgery <- 0
final_dataset$laser_glaucoma_surgery <- 0
final_dataset$surgery_count <- 0 
final_dataset$first_surgery_date <- NA

## now we are going to loop for the dataset and check for surgeries

## first we need to note our surgery codes
incisional_glaucoma_surgery <- c("66160","66170","66172","66179","66180","66185")
min_invasive_glaucoma_surgery <- c("66174","66175","66183","65850","65820","0191T","0449T")
laser_glaucoma_surgery <- c("66710","66711","65855")

## now we loop, patient_num represents the row of the final dataset we are in
for(patient_num in 1:nrow(final_dataset)){
  MRN <- final_dataset$MRN[patient_num] #pulls the patient ID
  patient_surgeries <- surgeries[MRN == surgeries$MRN,] #Pulls all surgeries associated with this patient id
  if(nrow(patient_surgeries) > 0) #checks to see if there were any surgeries
  {
    for(surgery_num in 1:nrow(patient_surgeries)) #goes through each surgery
    {
      if(is.element(patient_surgeries$real_cpt_code[surgery_num],incisional_glaucoma_surgery)) #checks for incisional surgery
      {
        final_dataset$any_glaucoma_surgey[patient_num] <- 1 
        final_dataset$incisional_glaucoma_surgery[patient_num] <- 1
        
        final_dataset$surgery_count[patient_num] <- 1 + final_dataset$surgery_count[patient_num]
        
        if(is.na(final_dataset$first_surgery_date[patient_num])) # checks to see if this is the patients first surgery, since the surgeries are time ordered we can assume this is the first surgery
        {
          final_dataset$first_surgery_date[patient_num] <- as.character(patient_surgeries$surgery_date[surgery_num])
          
        }
      }
      else if(is.element(patient_surgeries$real_cpt_code[surgery_num],min_invasive_glaucoma_surgery)) #checks for minimally invasive surgery
      {
        final_dataset$any_glaucoma_surgey[patient_num] <- 1
        final_dataset$min_invasive_glaucoma_surgery[patient_num] <- 1
        
        final_dataset$surgery_count[patient_num] <- 1 + final_dataset$surgery_count[patient_num]
        
        if(is.na(final_dataset$first_surgery_date[patient_num]))
        {
          final_dataset$first_surgery_date[patient_num] <- as.character(patient_surgeries$surgery_date[surgery_num])
        
        }
      }
      else if(is.element(patient_surgeries$real_cpt_code[surgery_num],laser_glaucoma_surgery)) #checks for laser surgery
      {
        final_dataset$any_glaucoma_surgey[patient_num] <- 1
        final_dataset$laser_glaucoma_surgery[patient_num] <- 1
        
         final_dataset$surgery_count[patient_num] <- 1 + final_dataset$surgery_count[patient_num]
        
        if(is.na(final_dataset$first_surgery_date[patient_num]))
        {
          final_dataset$first_surgery_date[patient_num] <- as.character(patient_surgeries$surgery_date[surgery_num])
         
        }
      }
    }
  }
}
  
## Since many of these variables are binaries, we will factor them so that R recognizes them as categorical versus numerical
final_dataset$any_glaucoma_surgey <- factor(final_dataset$any_glaucoma_surgey, levels = c(0,1), labels = c("No","Yes"))
final_dataset$min_invasive_glaucoma_surgery <- factor(final_dataset$min_invasive_glaucoma_surgery, levels = c(0,1), labels = c("No","Yes"))
final_dataset$laser_glaucoma_surgery <- factor(final_dataset$laser_glaucoma_surgery, levels = c(0,1), labels = c("No","Yes"))
final_dataset$incisional_glaucoma_surgery <- factor(final_dataset$incisional_glaucoma_surgery, levels = c(0,1), labels = c("No","Yes"))

```

## Vitals

So now we want to pull some health indicators we have from the vitals table!

```{r Vitals}

#first let us define the variables we want to create

final_dataset$systolic_max <- NA
final_dataset$systolic_min <- NA
final_dataset$systolic_mean <- NA
final_dataset$systolic_sd <- NA

final_dataset$diastolic_max <- NA
final_dataset$diastolic_min <- NA
final_dataset$diastolic_mean <- NA
final_dataset$diastolic_sd <- NA

final_dataset$BMI_mean <- NA
final_dataset$BMI_max <- NA
final_dataset$BMI_min <- NA

final_dataset$height <- NA
final_dataset$weight_min <- NA
final_dataset$weight_max <- NA
final_dataset$weight_mean <- NA

final_dataset$first_vitals_contact_date <- NA
final_dataset$vital_entries <- 0

final_dataset$pulse_min <- NA
final_dataset$pulse_max <- NA
final_dataset$pulse_mean <- NA

#need to recode the date variable because R can't interpret it correctly
vitals$contact_date <- as.Date(as.character.Date(vitals$contact_date),"%m/%d/%y")
#sort vitals by date
vitals <- vitals[order(vitals$contact_date),]
#converts height as coded into inches
vitals$height_inches <- sapply(strsplit(as.character(vitals$height),"'|\""), function(x){12*as.numeric(x[1]) + as.numeric(x[2])})

for(patient_num in 1:nrow(final_dataset)) #loop through all the patients
{
  MRN <- final_dataset$MRN[patient_num] #get their MRN
  surgery_date <- final_dataset$first_surgery_date[patient_num] #pull their surgery date 
  patient_vitals <- vitals[which(vitals$MRN == MRN),] #pull all their vital records

  #need to check if the patient had surgery, if they had surgery we will exclude all vital records which occurred after the date of first surgery
  if(!is.na(final_dataset$first_surgery_date[patient_num]))
    {patient_vitals <- patient_vitals[which(patient_vitals$contact_date < surgery_date),]}
 
  if(nrow(patient_vitals) > 0) #if there are any patient vitals, do the following
  {
    final_dataset$vital_entries[patient_num] <- nrow(patient_vitals)
    
    final_dataset$systolic_max[patient_num] <- max(patient_vitals$bp_systolic, na.rm = TRUE)
    final_dataset$systolic_min[patient_num] <- min(patient_vitals$bp_systolic, na.rm = TRUE)
    final_dataset$systolic_mean[patient_num] <- mean(patient_vitals$bp_systolic, na.rm = TRUE)
    final_dataset$systolic_sd[patient_num] <- sd(patient_vitals$bp_systolic, na.rm = TRUE)
    
    final_dataset$diastolic_max[patient_num] <- max(patient_vitals$bp_diastolic, na.rm = TRUE)
    final_dataset$diastolic_min[patient_num] <- min(patient_vitals$bp_diastolic, na.rm = TRUE)
    final_dataset$diastolic_mean[patient_num] <- mean(patient_vitals$bp_diastolic, na.rm = TRUE)
    final_dataset$diastolic_sd[patient_num] <- sd(patient_vitals$bp_diastolic, na.rm = TRUE)

    final_dataset$height[patient_num] <- mean(patient_vitals$height_inches, na.rm = TRUE)
    
    final_dataset$weight_min[patient_num] <- min(patient_vitals$Weight.in.lbs, na.rm = TRUE)
    final_dataset$weight_max[patient_num] <- max(patient_vitals$Weight.in.lbs, na.rm = TRUE)
    final_dataset$weight_mean[patient_num] <- mean(patient_vitals$Weight.in.lbs, na.rm = TRUE)
    
    final_dataset$BMI_max[patient_num] <- (final_dataset$weight_max[patient_num]*0.45)/((final_dataset$height[patient_num])*(final_dataset$height[patient_num])*.025*.025) 
    final_dataset$BMI_min[patient_num] <- (final_dataset$weight_min[patient_num]*0.45)/((final_dataset$height[patient_num])*(final_dataset$height[patient_num])*.025*.025) 
    final_dataset$BMI_mean[patient_num] <- (final_dataset$weight_mean[patient_num]*0.45)/((final_dataset$height[patient_num])*(final_dataset$height[patient_num])*.025*.025) 
    
    final_dataset$pulse_max[patient_num] <- max(patient_vitals$pulse, na.rm = TRUE) 
    final_dataset$pulse_min[patient_num] <- min(patient_vitals$pulse, na.rm = TRUE) 
    final_dataset$pulse_mean[patient_num] <- mean(patient_vitals$pulse, na.rm = TRUE) 
  
    final_dataset$first_vitals_contact_date[patient_num] <- as.character(patient_vitals$contact_date[1])
  }
    
}

```
### Social_Hx

Next, we will add information about alcohol and tobacco use

```{r social}

final_dataset$alcohol_use <- NA
final_dataset$smoking_status <- NA

for(patient_num in 1:nrow(final_dataset)){ # loop through all of the patients
  MRN <- final_dataset$MRN[patient_num] #pull MRN number
  patient_social <- social_hx[which(social_hx$MRN == MRN),] #pull their social_hx data
  
  if(is.element("Yes",patient_social$Alcohol_use)) #check for history of alcohol use
  {
    final_dataset$alcohol_use[patient_num] <- 1
  } else if(is.element("No",patient_social$Alcohol_use))
  {
    final_dataset$alcohol_use[patient_num] <- 0
  }
 
  if(is.element("Yes", patient_social$Tobacco_user) || is.element("Passive", patient_social$Tobacco_user)) #check for history of smoking
  {
    final_dataset$smoking_status[patient_num] <- 1
  } else if(is.element("Quit", patient_social$Tobacco_user))
  {
    final_dataset$smoking_status[patient_num] <- 2
  } else if(is.element("Never", patient_social$Tobacco_user))
  {
    final_dataset$smoking_status[patient_num] <- 0
  }
  
}
#since this data is categorical we need to recode it appropriately
final_dataset$smoking_status <- factor(final_dataset$smoking_status, levels = c(0,1,2), labels = c("Never", "Former","Current"))
final_dataset$alcohol_use <- factor(final_dataset$alcohol_use,levels=c(0,1), labels=c("No","Yes"))


```

## Diagnoses (ie comorbidities)

Here we will use diagnoses table data (ICD 9 or 10, will try both) to determine if patients exhibited specific comorbidities prior to their first surgery

```{r diagnoses}

# going to use the icd library

final_dataset$MI <- 0
final_dataset$CHF <- 0
final_dataset$PVD <- 0
final_dataset$Stroke <- 0
final_dataset$Dementia <- 0
final_dataset$Pulmonary <- 0
final_dataset$Rheumatic <- 0
final_dataset$PUD <- 0
final_dataset$LiverMild <- 0
final_dataset$DM <- 0
final_dataset$DMcx <- 0
final_dataset$Paralysis <- 0
final_dataset$Renal <- 0
final_dataset$Cancer <- 0
final_dataset$LiverSevere <- 0
final_dataset$Mets <- 0
final_dataset$HIV <- 0

#need to convert the the date value into a date that R recognizes
diagnoses$date <- as.Date(as.character.Date(diagnoses$contact_date),"%m/%d/%y")

for(patient_num in 1:nrow(final_dataset)) #loop through all patients
{
  MRN <- final_dataset$MRN[patient_num] #grab MRN number
  surgery_date <- final_dataset$first_surgery_date[patient_num] #pull surgery date
 
  patient_diagnoses <- diagnoses[which(diagnoses$MRN == MRN),] #pull al diagnoses data for said patient
  if(!is.na(surgery_date)) #if patient has had surgery, only include diagnoses preceding surgery
  {patient_diagnoses <- patient_diagnoses[which(patient_diagnoses$date < surgery_date),]}
  
  if(nrow(patient_diagnoses)>0) #if patients have any diagnosis records
  {
    #need to do appropriate formating for the icd library function icd_10_comorbi_charlson
    long_form_pd <- data.frame(matrix(nrow = nrow(patient_diagnoses), ncol = 0))
  
    long_form_pd$MRN <- MRN
    long_form_pd$icd10 <- patient_diagnoses$ICD10_code
    cc <- data.frame(icd10_comorbid_charlson(long_form_pd))
  
    if(cc$MI[1]) {final_dataset$MI[patient_num] <- 1} 
    if(cc$CHF[1]) {final_dataset$CHF[patient_num] <- 1}
    if(cc$PVD[1]) {final_dataset$PVD[patient_num] <- 1}
    if(cc$Stroke[1]) {final_dataset$Stroke[patient_num] <- 1}
    if(cc$Dementia[1]) {final_dataset$Dementia[patient_num] <- 1}
    if(cc$Pulmonary[1]) {final_dataset$Pulmonary[patient_num] <- 1}
    if(cc$Rheumatic[1]) {final_dataset$Rheumatic[patient_num] <- 1}
    if(cc$PUD[1]) {final_dataset$PUD[patient_num] <- 1}
    if(cc$LiverMild[1]) {final_dataset$LiverMild[patient_num] <- 1}
    if(cc$DM[1]) {final_dataset$DM[patient_num] <- 1}
    if(cc$DMcx[1]) {final_dataset$DMcx[patient_num] <- 1}
    if(cc$Paralysis[1]) {final_dataset$Paralysis[patient_num] <- 1}
    if(cc$Renal[1]) {final_dataset$Renal[patient_num] <- 1}
    if(cc$Cancer[1]) {final_dataset$Cancer[patient_num] <- 1}
    if(cc$LiverSevere[1]) {final_dataset$LiverSevere[patient_num] <- 1}
    if(cc$Mets[1]) {final_dataset$Mets[patient_num] <- 1}
    if(cc$HIV[1]) {final_dataset$HIV[patient_num] <- 1}
    
  }
  
  
}
#finally, given that this is categorical data, we need to code it appropriately
final_dataset$MI <- factor(final_dataset$MI, levels = c(0,1), labels = c("No","Yes"))
final_dataset$CHF <- factor(final_dataset$CHF, levels = c(0,1), labels = c("No","Yes"))
final_dataset$PVD <- factor(final_dataset$PVD, levels = c(0,1), labels = c("No","Yes"))
final_dataset$Stroke <- factor(final_dataset$Stroke, levels = c(0,1), labels = c("No","Yes"))
final_dataset$Dementia <- factor(final_dataset$Dementia, levels = c(0,1), labels = c("No","Yes"))
final_dataset$Pulmonary <- factor(final_dataset$Pulmonary, levels = c(0,1), labels = c("No","Yes"))
final_dataset$Rheumatic <- factor(final_dataset$Rheumatic, levels = c(0,1), labels = c("No","Yes"))
final_dataset$PUD <- factor(final_dataset$PUD, levels = c(0,1), labels = c("No","Yes"))
final_dataset$LiverMild <- factor(final_dataset$LiverMild, levels = c(0,1), labels = c("No","Yes"))
final_dataset$DM <- factor(final_dataset$DM, levels = c(0,1), labels = c("No","Yes"))
final_dataset$DMcx <- factor(final_dataset$DMcx, levels = c(0,1), labels = c("No","Yes"))
final_dataset$Paralysis <- factor(final_dataset$Paralysis, levels = c(0,1), labels = c("No","Yes"))
final_dataset$Renal <- factor(final_dataset$Renal, levels = c(0,1), labels = c("No","Yes"))
final_dataset$Cancer <- factor(final_dataset$Cancer, levels = c(0,1), labels = c("No","Yes"))
final_dataset$LiverSevere <- factor(final_dataset$LiverSevere, levels = c(0,1), labels = c("No","Yes"))
final_dataset$Mets <- factor(final_dataset$Mets, levels = c(0,1), labels = c("No","Yes"))
final_dataset$HIV <- factor(final_dataset$HIV, levels = c(0,1), labels = c("No","Yes"))


```

## Hospitalizations

Next we want to see how many hospitalizations PRECEDED surgery, if surgery occurred

```{r Hospitalizations}

final_dataset$ever_hospitalized <- 0
final_dataset$days_hospitalized <- 0
#need to convert the date value into one that R recognizes
hospitalizations$date <- as.Date(as.character.Date(hospitalizations$hosp_admsn_time),"%m/%d/%y")

for(patient_num in 1:nrow(final_dataset)) #loop through all patients
{
  MRN <- final_dataset$MRN[patient_num] #get patient MRN
  surgery_date <- final_dataset$first_surgery_date[patient_num] #get their first surgery data
  patient_hospitalizations <- hospitalizations[which(hospitalizations$MRN == MRN),] #pull their hospitalization records
  if(!is.na(surgery_date)) #if they had surgery, only include hospitalizations prior to first surgery
  {
    patient_hospitalizations <- patient_hospitalizations[which(patient_hospitalizations$date < surgery_date),]
  }
  if(nrow(patient_hospitalizations) > 0)
  {
    final_dataset$ever_hospitalized[patient_num] <- 1
    final_dataset$days_hospitalized[patient_num] <- length(unique(patient_hospitalizations$date))
  }
}
#recode this categorical variable
final_dataset$ever_hospitalized <- factor(final_dataset$ever_hospitalized, levels = c(0,1), labels = c("No","Yes"))

```

## Medications

Next, we want to add relevant information about medications prescribed and taken PRECEDING SURGERY EVENTS

```{r Medications}


#we want to only include medications that were used for at least two weeks so we need the duration of prescription (using Intent To Treat paradigm)
medications$start_med <- as.Date(as.character.Date(medications$start_date),"%m/%d/%y")
medications$end_med <- as.Date(as.character.Date(medications$end_date),"%m/%d/%y")
medications$med_duration <- difftime(medications$end_med, medications$start_med, units = "days")

# first we need to reformat the mediation table a little bit to incude the medication class as specificed in the medication class table
medications$med_class <- NA

for(i in 1:nrow(medications)) #loop through all of the medications
{
  
  med_id <- medications$medication_id[i] #pull medication ID
  med_row <- medication_class[medication_class$MEDICATION_ID == med_id,] #get the associated data from medication_class table for this medication
  med_class <- med_row[1,6] #get the name of the class
  medications$med_class[i] <- as.character(med_class) #insert name of class into the medications table
  
}
#so first we are going to just factor the medication classes, since it is categorical
medications$med_class <- factor(medications$med_class)

#then, we want a binary variable (YES/NO) for each medication class, so we are going to use the to.dummy functon to spread the single colum med_class into a set of binary columns

spread <- to.dummy(medications$med_class, "med_class")

#now we will bind all of these columns to the medication table so that the binary values are in the medication table
medications <- cbind(medications, spread)
##21:108

# we will use this df to track a single patients medications
medications_df <- data.frame(matrix(nrow = 0, ncol = 88))

for(patient_num in 1:nrow(final_dataset)) # loop through all of the patients
{
  MRN <- final_dataset$MRN[patient_num] #get their MRN
  surgery_date <- final_dataset$first_surgery_date[patient_num] # get their surgery date
  
  patient_meds <- medications[which(medications$MRN == MRN),] #get their medications records
 
  if(!is.na(surgery_date)) #if they had surgery, only include records prior to first surgery
  {
    patient_meds <- patient_meds[which(patient_meds$end_med < surgery_date),]
  }
  if(nrow(patient_meds) > 0) { #only include records in which medication was prescribed for at least two weeks
    patient_meds <- patient_meds[which(patient_meds$med_duration > 13),]
  }

  if(nrow(patient_meds) > 0) 
  {
    patient_sums <- colSums(patient_meds[,21:108]) #this will sum all of the binary columns, so that any column (medication class) with value greater than0 represents a medication class they took
    for(count in 1:length(patient_sums)) # loop through all of the results
    {
     
      if(patient_sums[count] > 0) #if value is greater than 0, recode it as 1 for consistency
      {
        patient_sums[count] <- 1
      }
    }
    medications_df <- rbind(medications_df, patient_sums) #add this to the table
  } else 
  {
    zeros <- integer(88) #this is if a patient has no medication records, we still need to input a bunch of zeros
    medications_df <- rbind(medications_df, zeros)
  }
}
#add teh appropriate column names
colnames(medications_df) <- colnames(spread)

#loop through and factor all as they are categorical variables
for(i in 1:ncol(medications_df))
{
  medications_df[,i] <- factor(medications_df[,i], levels = c(0,1),labels=c("No","Yes"))
}
#add the table of medications to the final dataset
final_dataset <- cbind(final_dataset,medications_df)


```
## Labs 

```{r Labs}

final_dataset$sodium <- NA
final_dataset$potassium <- NA
final_dataset$phos <- NA
final_dataset$bicarbonate <- NA
final_dataset$aniongap <- NA
final_dataset$creatinine <- NA
final_dataset$albumin <- NA
final_dataset$hgb <- NA
final_dataset$wbc <- NA
final_dataset$plt <- NA
final_dataset$iron <- NA
final_dataset$ironsaturation <- NA
final_dataset$mcv <- NA
final_dataset$rdw <- NA
final_dataset$mchc <- NA
final_dataset$ldlchol <- NA
final_dataset$triglycerides <- NA
final_dataset$nonhdlchol <- NA
final_dataset$hdl <- NA
final_dataset$glucose <- NA
final_dataset$a1c <- NA
final_dataset$tsh <- NA
final_dataset$sedrate <- NA
final_dataset$lactate <- NA


#Need to convert the data to a format R can read
labs$date <- as.Date(as.character.Date(labs$result_time),"%m/%d/%y")
for(patient_num in 1:nrow(final_dataset)) #loop through all patients
{
  MRN <- final_dataset$MRN[patient_num] #get patient MRN
  surgery_date <- final_dataset$first_surgery_date[patient_num] #get first surgery date
  patient_labs <- labs[which(labs$MRN == MRN),] #get all lab records
 
  if(!is.na(surgery_date)) #only include lab records prior to first surgery
  {
    patient_labs <- patient_labs[which(patient_labs$date < surgery_date),]
  }
  if(nrow(patient_labs) >= 0)
  {
    
    #SODIUM
    patient_sodium <- patient_labs[which(patient_labs$component_id == 2106),]
    final_dataset$sodium[patient_num] <- mean(patient_sodium$Value.as.Number, na.rm = TRUE)
    
    #POTASSIUM 
    patient_potassium <- patient_labs[which(patient_labs$component_id == 2107),]
    final_dataset$potassium[patient_num] <- mean(patient_potassium$Value.as.Number, na.rm = TRUE)
#PHOS
    patient_phos <- patient_labs[which(patient_labs$component_id == 2117),]
    final_dataset$phos[patient_num] <- mean(patient_phos$Value.as.Number, na.rm = TRUE)
#CALCIUM -- SAMPL TOO LOW 
   
#BICARBONATE
    patient_bicarbonate <- patient_labs[which(patient_labs$component_id == 2109),]
    final_dataset$bicarbonate[patient_num] <- mean(patient_bicarbonate$Value.as.Number, na.rm = TRUE)
#ANION GAP
    patient_aniongap <- patient_labs[which(patient_labs$component_id == 20835),]
    final_dataset$aniongap[patient_num] <- mean(patient_aniongap$Value.as.Number, na.rm = TRUE)
#CREATININE
    patient_creatinine <- patient_labs[which(patient_labs$component_id == 2104),]
    final_dataset$creatinine[patient_num] <- mean(patient_creatinine$Value.as.Number, na.rm = TRUE)
#ALBUMIN
    patient_albumin <- patient_labs[which(patient_labs$component_id == 2121),]
    final_dataset$albumin[patient_num] <- mean(patient_albumin$Value.as.Number, na.rm = TRUE)
#HGB 
    patient_hgb <- patient_labs[which(patient_labs$component_id == 7103),]
    final_dataset$hgb[patient_num] <- mean(patient_hgb$Value.as.Number, na.rm = TRUE)
#WBC
    patient_wbc <- patient_labs[which(patient_labs$component_id == 7100),]
    final_dataset$wbc[patient_num] <- mean(patient_wbc$Value.as.Number, na.rm = TRUE)
#PLT
    patient_plt <- patient_labs[which(patient_labs$component_id == 7111),]
    final_dataset$plt[patient_num] <- mean(patient_plt$Value.as.Number, na.rm = TRUE)
#IRON
    patient_iron <- patient_labs[which(patient_labs$component_id == 2160),]
    final_dataset$iron[patient_num] <- mean(patient_iron$Value.as.Number, na.rm = TRUE)
#IRON SATURATION
    patient_ironsaturation <- patient_labs[which(patient_labs$component_id == 2162),]
    final_dataset$ironsaturation[patient_num] <- mean(patient_ironsaturation$Value.as.Number, na.rm = TRUE)
#MCV
    patient_mcv <- patient_labs[which(patient_labs$component_id == 7105),]
    final_dataset$mcv[patient_num] <- mean(patient_mcv$Value.as.Number, na.rm = TRUE)
#RDW
    patient_rdw <- patient_labs[which(patient_labs$component_id == 7108),]
    final_dataset$rdw[patient_num] <- mean(patient_rdw$Value.as.Number, na.rm = TRUE)
#MCHC 
    patient_mchc <- patient_labs[which(patient_labs$component_id == 7107),]
    final_dataset$mchc[patient_num] <- mean(patient_mchc$Value.as.Number, na.rm = TRUE)
#LDL-CHOL (CALC)
    patient_ldlchol <- patient_labs[which(patient_labs$component_id == 2152),]
    final_dataset$ldlchol[patient_num] <- mean(patient_ldlchol$Value.as.Number, na.rm = TRUE)
#TRIGLYCERIDES 
    patient_triglycerides <- patient_labs[which(patient_labs$component_id == 2155),]
    final_dataset$triglycerides[patient_num] <- mean(patient_triglycerides$Value.as.Number, na.rm = TRUE)
#NON-HDL CHOLESTEROL 
    patient_nonhdlchol <- patient_labs[which(patient_labs$component_id == 20609),]
    final_dataset$nonhdlchol[patient_num] <- mean(patient_nonhdlchol$Value.as.Number, na.rm = TRUE)
#HDL 
    patient_hdl <- patient_labs[which(patient_labs$component_id == 2151),]
    final_dataset$hdl[patient_num] <- mean(patient_hdl$Value.as.Number, na.rm = TRUE)
#GLUCOSE 
    patient_glucose <- patient_labs[which(patient_labs$component_id == 2101),]
    final_dataset$glucose[patient_num] <- mean(patient_glucose$Value.as.Number, na.rm = TRUE)
#A1C 
    patient_a1c <- patient_labs[which(patient_labs$component_id == 2356),]
    final_dataset$a1c[patient_num] <- mean(patient_a1c$Value.as.Number, na.rm = TRUE)
#TSH
    patient_tsh <- patient_labs[which(patient_labs$component_id == 2195),]
    final_dataset$tsh[patient_num] <- mean(patient_tsh$Value.as.Number, na.rm = TRUE)
#SED RATE 
    patient_sedrate <- patient_labs[which(patient_labs$component_id == 7200),]
    final_dataset$sedrate[patient_num] <- mean(patient_sedrate$Value.as.Number, na.rm = TRUE)
#LACTATE
    patient_lactate <- patient_labs[which(patient_labs$component_id == 2175),]
    final_dataset$lactate[patient_num] <- mean(patient_lactate$Value.as.Number, na.rm = TRUE)
  }


}

```

## Output the Dataset

```{r output}
# get rid of the patient MRNS
final_dataset$MRN <- NA

#output the final dataset
write.csv(final_dataset,"Dec8_Dataset.csv")

```





Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
