---
title: "K-Cross Validation Artificial NN"
author: "Charles Marks"
date: "December 8, 2018"
output: html_document
---

Got most of this stuff from https://www.analyticsvidhya.com/blog/2017/09/creating-visualizing-neural-network-in-r/

```{r global options}
library(tidyverse)
library(tableone)
library(PerformanceAnalytics)
library(psych)
library(ROCR)
library(randomForest)
library(neuralnet)
options(digits = 2)
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```
## Creating the five groups for cross-validation

So, like with the logistic regression model, we will assign the participants to 5 groups to do 80-20 cross validation and choose the best performing model.  For this doc, we have excluded the code for loading the data and preparing it for analysis.  See the kcrossval.rmd for that code.

## Creating the five groups for cross-validation

So, unlike with log-regression and random forests, ANN requires numerical data, so before sorting into random groups we need to make the data numerical and then we need to scale it.  So, first we will scale the data and then we shall assign to random groups.

## Defining Our Five Groups

```{r define groups}

#first scale the data
numeric_data <- clean_data

for(i in 1:ncol(numeric_data)) # this loop goes through all of the variables and makes the non-numeric data numeric
{
  if(!is.numeric(numeric_data[,i]))
  {
    numeric_data[,i] <- as.numeric(numeric_data[,i])
  }
  
  
}

# these three lines scale all of the data frame
max <- apply(numeric_data,2,max)
min <- apply(numeric_data,2,min)
numeric_data <- as.data.frame(scale(numeric_data, center = min, scale = max - min))

# now lets do our randomization!
values <- rep(1:5, length.out=391)
random <- sample(values)
numeric_data$cv_group <- numbers

# finally, we create our 5 training and test data sets based on their assigned random groups
train.one <- numeric_data[which(numeric_data$cv_group!= 1),]
test.one <- numeric_data[which(numeric_data$cv_group== 1),]

train.two <- numeric_data[which(numeric_data$cv_group!= 2),]
test.two <- numeric_data[which(numeric_data$cv_group== 2),]

train.three <- numeric_data[which(numeric_data$cv_group!= 3),]
test.three <- numeric_data[which(numeric_data$cv_group== 3),]

train.four <- numeric_data[which(numeric_data$cv_group!= 4),]
test.four <- numeric_data[which(numeric_data$cv_group== 4),]

train.five <- numeric_data[which(numeric_data$cv_group!= 5),]
test.five <- numeric_data[which(numeric_data$cv_group== 5),]

```

## Running Cross Val

Just as in our previous models, we will run the model and then present AUCs to pick the model with the best fit

### Test 1

```{r test 1, include=FALSE}

set.seed(2)

model.ann.one <- neuralnet(any_glaucoma_surgey ~ Age + Gender +systolic_max + systolic_min + systolic_mean  + diastolic_max + diastolic_min + diastolic_mean +pulse_min + pulse_max + pulse_mean  +ever_hospitalized + days_hospitalized + CHF + PVD + Stroke + Dementia + Pulmonary + LiverMild + DM + DMcx  + Renal + Cancer + Mets + med_class.Analgesics_._non.opioids + med_class.Analgesics_._opioids  +  med_class.Anti.rheumatic +  med_class.Antianxiety_agents  + med_class.Antiasthmatic + med_class.Anticoagulants+med_class.Anticonvulsant+med_class.Antidepressants+med_class.Antidiabetic+med_class.Antiemetics+med_class.Antihyperlipidemic+med_class.Antihypertensive+med_class.Beta_blockers+med_class.Calcium_blockers+med_class.Corticosteroids+med_class.Cough.Cold+med_class.Decongestants+med_class.Dermatological+med_class.Diuretics+med_class.Laxatives+med_class.Macrolide_antibiotics+med_class.Misc._antiinfectives+med_class.Ophthalmic+med_class.Ulcer_drugs, 
                 data=train.one, hidden = 2
                 )
# so, this next line, we need a dataframe that only contains the variables included in the model, these numbers represent the columns of the data frame of all the variables contained in the model...it's not pretty but it works
test.one.cv <- test.one[,c(3,4,13,14,15,17,18,19,30,31,32,36,37,38,39,40,43,44,45,47,48,50,52,53,55,56,61,63,65,66,67,68,69,72,75,76,87,88,94,95,96,97,101,108,110,115,126,136)]

### this line is for reproduction (talk to developer about why this is necessary)test.one.cv <- test.one[,c(4,5,14,15,16,18,19,20,31,32,33,37,38,39,40,41,44,45,46,48,49,51,53,54,56,57,62,64,66,67,68,69,70,73,76,77,88,89,95,96,97,98,102,109,111,116,127,137)]

prob.one <- compute(model.ann.one,test.one.cv)
prob.one.results <- prob.one$net.result

pred.one <- ROCR::prediction(prob.one.results,test.one$any_glaucoma_surgey)
perf.one <- ROCR::performance(pred.one,"tpr","fpr")
plot(perf.one)

nn.one.auc <- performance(pred.one,"auc")@y.values[[1]]
```

### Test 2

```{r test 2, include=FALSE}
set.seed(2)

model.ann.two <- neuralnet(any_glaucoma_surgey ~ Age + Gender +systolic_max + systolic_min + systolic_mean  + diastolic_max + diastolic_min + diastolic_mean +pulse_min + pulse_max + pulse_mean  +ever_hospitalized + days_hospitalized + CHF + PVD + Stroke + Dementia + Pulmonary + LiverMild + DM + DMcx  + Renal + Cancer + Mets + med_class.Analgesics_._non.opioids + med_class.Analgesics_._opioids  +  med_class.Anti.rheumatic +  med_class.Antianxiety_agents  + med_class.Antiasthmatic + med_class.Anticoagulants+med_class.Anticonvulsant+med_class.Antidepressants+med_class.Antidiabetic+med_class.Antiemetics+med_class.Antihyperlipidemic+med_class.Antihypertensive+med_class.Beta_blockers+med_class.Calcium_blockers+med_class.Corticosteroids+med_class.Cough.Cold+med_class.Decongestants+med_class.Dermatological+med_class.Diuretics+med_class.Laxatives+med_class.Macrolide_antibiotics+med_class.Misc._antiinfectives+med_class.Ophthalmic+med_class.Ulcer_drugs, 
                 data=train.two, hidden = 2
                 )

test.two.cv <- test.two[,c(3,4,13,14,15,17,18,19,30,31,32,36,37,38,39,40,43,44,45,47,48,50,52,53,55,56,61,63,65,66,67,68,69,72,75,76,87,88,94,95,96,97,101,108,110,115,126,136)]
##test.two.cv <- test.two[,c(4,5,14,15,16,18,19,20,31,32,33,37,38,39,40,41,44,45,46,48,49,51,53,54,56,57,62,64,66,67,68,69,70,73,76,77,88,89,95,96,97,98,102,109,111,116,127,137)]
prob.two <- compute(model.ann.two,test.two.cv)
prob.two.results <- prob.two$net.result

pred.two <- ROCR::prediction(prob.two.results,test.two$any_glaucoma_surgey)
perf.two <- ROCR::performance(pred.two,"tpr","fpr")
plot(perf.two)

nn.two.auc <- performance(pred.two,"auc")@y.values[[1]]
```

### Test 3

```{r test 3, include=FALSE}


set.seed(2)

model.ann.three <- neuralnet(any_glaucoma_surgey ~ Age + Gender +systolic_max + systolic_min + systolic_mean  + diastolic_max + diastolic_min + diastolic_mean +pulse_min + pulse_max + pulse_mean  +ever_hospitalized + days_hospitalized + CHF + PVD + Stroke + Dementia + Pulmonary + LiverMild + DM + DMcx  + Renal + Cancer + Mets + med_class.Analgesics_._non.opioids + med_class.Analgesics_._opioids  +  med_class.Anti.rheumatic +  med_class.Antianxiety_agents  + med_class.Antiasthmatic + med_class.Anticoagulants+med_class.Anticonvulsant+med_class.Antidepressants+med_class.Antidiabetic+med_class.Antiemetics+med_class.Antihyperlipidemic+med_class.Antihypertensive+med_class.Beta_blockers+med_class.Calcium_blockers+med_class.Corticosteroids+med_class.Cough.Cold+med_class.Decongestants+med_class.Dermatological+med_class.Diuretics+med_class.Laxatives+med_class.Macrolide_antibiotics+med_class.Misc._antiinfectives+med_class.Ophthalmic+med_class.Ulcer_drugs, 
                 data=train.three, hidden = 2
                 )

test.three.cv <- test.three[,c(3,4,13,14,15,17,18,19,30,31,32,36,37,38,39,40,43,44,45,47,48,50,52,53,55,56,61,63,65,66,67,68,69,72,75,76,87,88,94,95,96,97,101,108,110,115,126,136)]
##test.three.cv <- test.three[,c(4,5,14,15,16,18,19,20,31,32,33,37,38,39,40,41,44,45,46,48,49,51,53,54,56,57,62,64,66,67,68,69,70,73,76,77,88,89,95,96,97,98,102,109,111,116,127,137)]
prob.three <- neuralnet::compute(model.ann.three,test.three.cv)
prob.three.results <- prob.three$net.result

pred.three <- ROCR::prediction(prob.three.results,test.three$any_glaucoma_surgey)
perf.three <- ROCR::performance(pred.three,"tpr","fpr")
plot(perf.three)

nn.three.auc <- performance(pred.three,"auc")@y.values[[1]]
```

```{r get sensitivity and specificity}
opt.cut = function(perf, pred){
    cut.ind = mapply(FUN=function(x, y, p){
        d = (x - 0)^2 + (y-1)^2
        ind = which(d == min(d))
        c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
            cutoff = p[[ind]])
    }, perf@x.values, perf@y.values, pred@cutoffs)
}
print(opt.cut(perf.two, pred.two))


acc.perf = performance(pred.two, measure = "acc")
plot(acc.perf)


ind = which.max( slot(acc.perf, "y.values")[[1]] )
acc = slot(acc.perf, "y.values")[[1]][ind]
cutoff = slot(acc.perf, "x.values")[[1]][ind]
print(c(accuracy= acc, cutoff = cutoff))

```
### Test 4

```{r test 4, include=FALSE}

set.seed(2)

model.ann.four <- neuralnet(any_glaucoma_surgey ~ Age + Gender +systolic_max + systolic_min + systolic_mean  + diastolic_max + diastolic_min + diastolic_mean +pulse_min + pulse_max + pulse_mean  +ever_hospitalized + days_hospitalized + CHF + PVD + Stroke + Dementia + Pulmonary + LiverMild + DM + DMcx  + Renal + Cancer + Mets + med_class.Analgesics_._non.opioids + med_class.Analgesics_._opioids  +  med_class.Anti.rheumatic +  med_class.Antianxiety_agents  + med_class.Antiasthmatic + med_class.Anticoagulants+med_class.Anticonvulsant+med_class.Antidepressants+med_class.Antidiabetic+med_class.Antiemetics+med_class.Antihyperlipidemic+med_class.Antihypertensive+med_class.Beta_blockers+med_class.Calcium_blockers+med_class.Corticosteroids+med_class.Cough.Cold+med_class.Decongestants+med_class.Dermatological+med_class.Diuretics+med_class.Laxatives+med_class.Macrolide_antibiotics+med_class.Misc._antiinfectives+med_class.Ophthalmic+med_class.Ulcer_drugs, 
                 data=train.four, hidden = 2
                 )

test.four.cv <- test.four[,c(3,4,13,14,15,17,18,19,30,31,32,36,37,38,39,40,43,44,45,47,48,50,52,53,55,56,61,63,65,66,67,68,69,72,75,76,87,88,94,95,96,97,101,108,110,115,126,136)]
prob.four <- compute(model.ann.four,test.four.cv)
prob.four.results <- prob.four$net.result

pred.four <- ROCR::prediction(prob.four.results,test.four$any_glaucoma_surgey)
perf.four <- ROCR::performance(pred.four,"tpr","fpr")
plot(perf.four)

nn.four.auc <- performance(pred.four,"auc")@y.values[[1]]
```
### Test 5

```{r test 5, include=FALSE}

set.seed(2)

model.ann.five <- neuralnet(any_glaucoma_surgey ~ Age + Gender +systolic_max + systolic_min + systolic_mean  + diastolic_max + diastolic_min + diastolic_mean +pulse_min + pulse_max + pulse_mean  +ever_hospitalized + days_hospitalized + CHF + PVD + Stroke + Dementia + Pulmonary + LiverMild + DM + DMcx  + Renal + Cancer + Mets + med_class.Analgesics_._non.opioids + med_class.Analgesics_._opioids  +  med_class.Anti.rheumatic +  med_class.Antianxiety_agents  + med_class.Antiasthmatic + med_class.Anticoagulants+med_class.Anticonvulsant+med_class.Antidepressants+med_class.Antidiabetic+med_class.Antiemetics+med_class.Antihyperlipidemic+med_class.Antihypertensive+med_class.Beta_blockers+med_class.Calcium_blockers+med_class.Corticosteroids+med_class.Cough.Cold+med_class.Decongestants+med_class.Dermatological+med_class.Diuretics+med_class.Laxatives+med_class.Macrolide_antibiotics+med_class.Misc._antiinfectives+med_class.Ophthalmic+med_class.Ulcer_drugs, 
                 data=train.five, hidden = 2
                 )

test.five.cv <- test.five[,c(3,4,13,14,15,17,18,19,30,31,32,36,37,38,39,40,43,44,45,47,48,50,52,53,55,56,61,63,65,66,67,68,69,72,75,76,87,88,94,95,96,97,101,108,110,115,126,136)]
prob.five <- compute(model.ann.five,test.five.cv)
prob.five.results <- prob.five$net.result

pred.five <- ROCR::prediction(prob.five.results,test.five$any_glaucoma_surgey)
perf.five <- ROCR::performance(pred.five,"tpr","fpr")
plot(perf.five)

nn.five.auc <- performance(pred.five,"auc")@y.values[[1]]

```
### COMAPRE ROC

This prints out the AUROC scores for comparison

```{r roc comapre}

print(nn.one.auc)
print(nn.two.auc)
print(nn.three.auc)
print(nn.four.auc)
print(nn.five.auc)

```

## Results of Best Model

```{r looking at the best}

prob.two.transform <- (prob.two.results*(max(train.two$any_glaucoma_surgey) - min(train.two$any_glaucoma_surgey))) + min(train.two$any_glaucoma_surgey)


plot(test.two$any_glaucoma_surgey,prob.two.transform)
```

## Evaluating This Model

### Discrimination (C-Statistic aka AUC)

```{r AUC}

plot(perf.two,main="ROC Curve for Artificial Neural Network",col=2,lwd=2)
abline(a=0,b=1,lwd=2,lty=2,col="gray")


```

### Caibration (Hoslem-Lemeshow)

```{r calibration}


library(gbm)
calibrate.plot(test.two$any_glaucoma_surgey, prob.two.results)
```

## Saving the final docs!

```{r Saving Docs}

clean_data$cv_group <- numeric_data$cv_group

final_test_set <- clean_data[which(clean_data$cv_group!= 2),]
final_training_set <- clean_data[which(clean_data$cv_group== 2),]


#importance_scores <- importance(model.rf.two)

write.csv(final_test_set, "Final_Test_Set_ANN_Dec10.csv")
write.csv(final_training_set, "Final_Training_Set_ANN_Dec10.csv")


```
